---
title: "lab14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Credit dataset

```{r}
library(ISLR)
names(Credit)
summary(Credit)
M = lm(Balance ~ . - ID, data=Credit)
```


# LASSO with glmnet

```{r}
library(glmnet)
X = scale(model.matrix(M)[,-1], center=TRUE, scale=TRUE)
Y = Credit$Balance
cvG = cv.glmnet(X, Y)
plot(cvG)
```

# Coefficient path

```{r}
G = glmnet(X, Y)
plot(G, xvar='lambda')
abline(v=log(cvG$lambda.min), col='green', lty=2)
abline(v=log(cvG$lambda.1se), col='green', lty=2)
```

# Extracting a coefficient

```{r getcoef}
coef(G, s=cvG$lambda.1se)
```

# Dangers of using LASSO selected variables in inference

Let's generate some null data and then compute p-values and confidence intervals.
If p-values are valid, they should be uniformly distributed. If confidence
intervals are valid, 95% should cover 0.
```{r}
simulate_null = function(n=200, p=100) {
  while (TRUE) {
  X = matrix(rnorm(n*p), n, p)
  Y = rnorm(n)
  cvG = cv.glmnet(X, Y)
  G = glmnet(X, Y)
  beta.lasso = as.numeric(coef(G, s=cvG$lambda.min)[-1])
  if (sum(beta.lasso != 0) > 1) { # 1 instead of 0 to avoid silly matrix shaping
    selected = beta.lasso != 0
    M = lm(Y ~ X[,selected])
    intervals = confint(M)[-1,]
    covered = (intervals[,1] < 0) * (intervals[,2] > 0)
    pvalues = summary(M)$coef[,4][-1]
    return(list(covered=covered, pvalues=pvalues))
    break
    }
  }
}
```

Let's run this several times.
```{r}
coverage = c()
pvalues = c()
for (i in 1:20) {
  results = simulate_null()
  coverage = c(coverage, results$covered)
  pvalues = c(pvalues, results$pvalues)
}
print(mean(coverage))
print(mean(pvalues < 0.05))
plot(ecdf(pvalues))
```

Let's assume we have an additional validation data set.

```{r}
simulate_null_validation = function(n=200, p=100) {
  while (TRUE) {
  X = matrix(rnorm(n*p), n, p)
  Y = rnorm(n)
  cvG = cv.glmnet(X, Y)
  G = glmnet(X, Y)
  beta.lasso = as.numeric(coef(G, s=cvG$lambda.min)[-1])
  if (sum(beta.lasso != 0) > 1) { # 1 instead of 0 to avoid silly matrix shaping
    selected = beta.lasso != 0
    Xvalid = matrix(rnorm(n*p), n, p)
    Yvalid = rnorm(n)
    M = lm(Yvalid ~ Xvalid[,selected])
    intervals = confint(M)[-1,]
    covered = (intervals[,1] < 0) * (intervals[,2] > 0)
    pvalues = summary(M)$coef[,4][-1]
    return(list(covered=covered, pvalues=pvalues))
    break
    }
  }
}

coverage_valid = c()
pvalues_valid = c()
for (i in 1:20) {
  results = simulate_null_validation()
  coverage_valid = c(coverage_valid, results$covered)
  pvalues_valid = c(pvalues_valid, results$pvalues)
}
print(mean(coverage_valid))
print(mean(pvalues_valid < 0.05))
plot(ecdf(pvalues_valid))
```

# Principal components regression

```{r}
set.seed(2)
pcr.fit = pcr(Salary ~ ., data=Hitters, scale=TRUE, validation="CV")
summary(pcr.fit)
validationplot(pcr.fit, val.type="MSEP")
pcr.fit.selecte = pcr(Salary ~ ., data=Hitters, scale=TRUE, validation="CV", ncomp=7)
```