---
output:
  html_document: default
  pdf_document: default
---
```{r}
library(ggplot2)
library(splines)
set.seed(1)
```

Define a true function $f$
```{r}
f = function(x) {
   x^2 - 0.2*x^2.3333
}
```
Now, we sample a random observation of the function at 10 input points with normal errors. We repeat this 4 times and show each repetition in different colors.

```{r}
d = data.frame("x"=seq(0,100,0.1))
d$f = f(d$x)
# Samples
s = data.frame("x"=c(3,10,80,70,65,43,90,40,32,15,98))
means = f(s$x)
s$f1 = rnorm(n=length(s$x),mean=means,sd=rep(250,length(s$x)))
s$f2 = rnorm(n=length(s$x),mean=means,sd=rep(250,length(s$x)))
s$f3 = rnorm(n=length(s$x),mean=means,sd=rep(250,length(s$x)))
s$f4 = rnorm(n=length(s$x),mean=means,sd=rep(250,length(s$x)))
lims = scale_y_continuous(limits=c(-500,1300),name="f(x)")
```

```{r}
ggplot(d,aes(x=x,y=f)) + geom_line()  + geom_point(data=s,aes(x=x,y=f1),color='red') + lims
```
```{r}
ggplot(d,aes(x=x,y=f)) + geom_line()  + geom_point(data=s,aes(x=x,y=f2),color='blue') + lims
```
```{r}
ggplot(d,aes(x=x,y=f)) + geom_line()  + geom_point(data=s,aes(x=x,y=f3),color='green') + lims
```
```{r}
ggplot(d,aes(x=x,y=f)) + geom_line()  + geom_point(data=s,aes(x=x,y=f4),color='orange')+ lims
```

Now, for each realization of our experiment, we fit a spline that estimates the true function in black.
```{r}
# Spline fits
fit = lm(f1~bs(x,df=5),data=s)
d$f1 = predict(object=fit,newdata=list(x=d$x))
fit = lm(f2~bs(x,df=5),data=s)
d$f2 = predict(object=fit,newdata=list(x=d$x))
fit = lm(f3~bs(x,df=5),data=s)
d$f3 = predict(object=fit,newdata=list(x=d$x))
fit = lm(f4~bs(x,df=5),data=s)
d$f4 = predict(object=fit,newdata=list(x=d$x))
```
```{r}
ggplot(d,aes(x=x,y=f)) + geom_line()  + geom_point(data=s,aes(x=x,y=f1),color='red') + geom_line(data=d,aes(x=x,y=f1),color='red') + lims
```
```{r}
ggplot(d,aes(x=x,y=f)) + geom_line()  + geom_point(data=s,aes(x=x,y=f2),color='blue') + geom_line(data=d,aes(x=x,y=f2),color='blue') + lims
```
```{r}
ggplot(d,aes(x=x,y=f)) + geom_line()  + geom_point(data=s,aes(x=x,y=f3),color='green') + geom_line(data=d,aes(x=x,y=f3),color='green') + lims
```
```{r}
ggplot(d,aes(x=x,y=f)) + geom_line()  + geom_point(data=s,aes(x=x,y=f4),color='orange') + geom_line(data=d,aes(x=x,y=f4),color='orange') + lims
```

Suppose we want to make a prediction at a specific point $x_0$. 
The procedure we've employed, fitting a spline, has a Mean Squared Error at this test point. 
We know that this is split into variance and squared bias components.
The variance measures the variability of $\hat{f}(x_0)$, the fit evaluated at the test point, when we repeat the sampling of the data (all 10 points). 

We can visualize this by examining the 4 fits above at ￼$x_0=70$￼ .
```{r}
valuesAt70 = data.frame('x'=rep(70,4),'f'=c(d$f1[701],d$f2[701],d$f3[701],d$f4[701])) 
ggplot(d,aes(x=x,y=f)) + geom_line() + geom_line(data=d,aes(x=x,y=f1),color='red') + geom_line(data=d,aes(x=x,y=f2),color='blue') + geom_line(data=d,aes(x=x,y=f3),color='green') + geom_line(data=d,aes(x=x,y=f4),color='orange') + geom_point(data=valuesAt70,aes(x=x,y=f)) + lims
```
If we repeated the experiment many times, we would be able to estimate the variance exactly.
The bias measures whether the estimates ￼ tend to err "in the same direction" " with respect to the true at $x_0=70$.
Just from 4 repetitions of the experiment, it is hard to tell. In this case the estimate fell below the true
function once, and above once at  $x_0=70$. 

If we were able to repeat the experiment many times, we could measure the bias by taking the difference between the average prediction, plotted below as a dashed line and the true function.
```{r}
d$average = (d$f1+d$f2+d$f3+d$f4)/4
valuesAt70 = data.frame(x=rep(70,2), f=c(d$f[701],d$average[701])) 
ggplot(d,aes(x=x,y=f)) + geom_line() + geom_line(data=d,aes(x=x,y=average),linetype='dashed') + geom_point(data=valuesAt70,aes(x=x,y=f)) + lims
```
From this limited number of repetitions, there seems to be almost no bias at $x_0=70$.

￼